Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-11 article/article.tex
PCA Components results added

\textbf{\emph{ROI mapping.}} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used. As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field. We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC'. These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}. To further reduce the number of features, average voxel values and standard deviations within these regions are taken as new input features. As other higher-order features such as $kurtosis$ and $skewness$ within ROIs did not increase predictability, these were not included in the subsequent analysis. The disadvantage of using ROIs is that they require \textit{a priori} knowledge. When this knowledge is inaccurate, classification accuracy may not be optimal.

@@

\begin{table*}[htpb]
\centering
{\small
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Extraction/Selection Method} & \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.46 &  - & - & - \\
M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
\hline
\end{tabular}}
\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure. \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.}
\label{tab:results}
\end{table*}

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which suggests that...

>>>>

\textbf{\emph{ROI mapping.}} ROI mapping is a form of feature selection, wherein only a subset of the N-dimensional input space is used. As described in the introduction, ROIs are anatomically defined brain areas classified as relevant by an expert in the field. We used 7 of the 25 defined ROIs, labeled  'CALC' 'LIPL' 'LT' 'LTRIA' 'LOPER' 'LIPS' and  'LDLPFC'. These ROIs have previously been found to be most relevant in discriminating between the cognitive states of looking at sentences or pictures \cite{wa:2003betw}. To further reduce the number of features, average voxel values within these regions are taken as new input features. As other higher-order features such as the standard deviation, kurtosis and skewness within ROIs did not increase predictability, these were not included in the subsequent analysis. The disadvantage of using ROIs is that they require \textit{a priori} knowledge. When this knowledge is inaccurate, classification accuracy may not be optimal.

@@

\begin{table*}[htpb]
\centering
{\small
\begin{tabular}{c|c|c|c|c}
& \textbf{Naive Bayes} & \textbf{Linear SVM} & \textbf{Naive Bayes*} & \textbf{Multivariate*} \\ \hline
PCA & 61.46 &  - & - & - \\
M ROIs & 70.21 & 72.5 & 74.79 & 76.67 \\
Norm-M ROIs & 71.67 & 71.25 & 75.83 & 77.08 \\
Norm-M-SD ROIs & 65.63 & 69.79 & 69.38 & 65.0 \\
Norm/Vox-M ROIs & 73.13 & 76.67 & 76.88 & 79.58 \\
\end{tabular}}
\caption{Performance results (in \%) of the three classifiers on different methods for feature selection and structure. \emph{M} stands for \emph{Mean}, \emph{Norm} for \emph{Normalisation}, \emph{SD} for \emph{Standard Deviation}, and \emph{Norm/Vox} for \emph{Normalisation per Voxel index}.}
\label{tab:results}
\end{table*}

@@

From Table~\ref{tab:results} we can distinguish the fact that when PCA is used for feature extraction, the performance of GNB is lower than when other feature selection methods are used. The confidence interval of the predictive performance of GNB when PCA is used is 95\% $CI: 56.9 - 65.8$ which is significantly worse than other dimension reduction techniques. This suggests that maximizing the variance of the data does not maximize information. Another explanation could be that the data is not suited to be aligned by voxel index (see section~\ref{sec:dimensionalityReduction}). The performance listed in Table~\ref{tab:results} is for PCA using 149 components. This value was found by performing classification with a varying amount of components, the results for which are listed in figure~\ref{fig:pca_components}.

\begin{figure}
	\centering
	\includegraphics[width=90mm]{figures/pca_plot_by_index.png}
  	\caption{Classification accuracy of GNB after performing PCA with a varying amount of components.}
  	\label{fig:pca_components}
\end{figure}
